\class : Problem Set \psetnum

### 2022-04-12 558_exoplanets/pset1/solutions.tex
fix units and whatnot

\question{\textbf{4. Fit planet parameters}}
\answer{
    Now for the main event! I fitted for the planet parameters $(P, k_*, t_p, \gamma, \omega, e)$ using \texttt{emcee}. I set the initial guess for $P$ using the answer from Q3. For $k_*$ I used $0.5(v_{rv, {\rm max}} - v_{rv, {\rm min}})$ from the definition. I guessed values for the other four just by playing around with them until the fit looked sort of close. This gave initial guesses as
    \begin{equation}
        \mathrm{initial\ guesses} = [111.47 \unit{days}, 0.4973 \unit{m}{s^{-1}}, 10.7 \unit{days}, -0.1 \unit{m}{s^{-1}}, 5 \unit{rad}, 0.92]
    \end{equation}
    for $P, k_*, t_p, \gamma, \omega, e$ respectively. I then applied the following bounds as a prior
    \begin{equation}
        \begin{split}
            \mathrm{bounds} =
                [&(110, 112) \unit{days}, (0.2, 0.7) \unit{m}{s^{-1}}, (10.2, 11.2) \unit{days},\\
                &(-1, 1) \unit{m}{s^{-1}}, (0.01, 2 \pi) \unit{rad}, (0, 1)]
        \end{split}
    \end{equation}
    Next I ran \texttt{emcee} using 16 walkers, with 5000 steps each and a 1000 step burn in. I've plotted the parameter distribution as a corner plot and \textbf{I put it at the end of this document}.

@@

    To get the best fit for the parameters, I take the median values from each of these distributions. The best fit values are:
    \begin{equation}
        \boxed{ P, k_*, t_p, \gamma, \omega, e = 111.31 \unit{days}, 0.471 \unit{m}{s^{-1}}, 10.92 \unit{days}, -0.0019 \unit{m}{s^{-1}}, 5.251 \unit{rad}, 0.932}
    \end{equation}
    I plot this with the folded data (on the same period as the fit) in the plot below - and it shows that we have good agreement, fun!
    \begin{center}
        \includegraphics[width=0.8\textwidth]{figures/best_fit.pdf}
    \end{center}
    So now I got intrigued and wondered what this meant for the mass of the planet. Since we know that $k_*$ is defined as
    \begin{equation}
        k_* = \qty(\frac{2 \pi G}{P (1 - e^2)})^{1/3} \frac{M_p \sin \iota}{(M_* + M_p)^{2/3}},
    \end{equation}
    where $\iota$ is the inclination and $M_*$ is the star mass. It should be fairly reasonable to rearrange this to find the planet mass and make the assumption that $M_* \gg M_p$ (this worries me slightly but I figure it'll be fine except in extreme cases). This gives that the planet mass is
    \begin{equation}
        M_p = \frac{k_* M_*^{2/3}}{\sin \iota} \qty(\frac{2 \pi G}{P (1 - e^2)})^{-1/3},
    \end{equation}
    So now we know everything in this equation from the best fit except $\iota$ and $M_*$. I made a grid of these values and plotted the planet mass as a contour plot below along with some lines indicating Earth and Neptune.

@@

    Eric hasn't given us any constraints on either of these parameters so it's hard to say more but if we could constrain the inclination to be larger than a small angle and had a guess for the stellar type then we'd quickly get a good idea for the planet mass from this plot! :D
}

>>>>

\question{\textbf{4. Fit planet parameters}}
\answer{
    Now for the main event! I fitted for the planet parameters $(P, k_*, t_p, \gamma, \omega, e)$ using \texttt{emcee}. I set the initial guess for $P$ using the answer from Q3. For $k_*$ I used $0.5(v_{rv, {\rm max}} - v_{rv, {\rm min}})$ from the definition. I guessed values for the other four just by playing around with them until the fit looked sort of close. This gave initial guesses as
    \begin{equation}
        \mathrm{initial\ guesses} = [111.47 \unit{days}, 0.4973 \unit{km}{s^{-1}}, 10.7 \unit{days}, -0.1 \unit{km}{s^{-1}}, 5 \unit{rad}, 0.92]
    \end{equation}
    for $P, k_*, t_p, \gamma, \omega, e$ respectively. I then applied the following bounds as a prior
    \begin{equation}
        \begin{split}
            \mathrm{bounds} =
                [&(110, 112) \unit{days}, (0.2, 0.7) \unit{km}{s^{-1}}, (10.2, 11.2) \unit{days},\\
                &(-1, 1) \unit{km}{s^{-1}}, (0.01, 2 \pi) \unit{rad}, (0, 1)]
        \end{split}
    \end{equation}
    Next I ran \texttt{emcee} using 16 walkers, with 5000 steps each and a 1000 step burn in. I've plotted the parameter distribution as a corner plot and \textbf{I put it at the end of this document}.

@@

    To get the best fit for the parameters, I take the median values from each of these distributions. The best fit values are:
    \begin{equation}
        \boxed{ P, k_*, t_p, \gamma, \omega, e = 111.31 \unit{days}, 0.471 \unit{km}{s^{-1}}, 10.92 \unit{days}, -0.0019 \unit{km}{s^{-1}}, 5.251 \unit{rad}, 0.932}
    \end{equation}
    I plot this with the folded data (on the same period as the fit) in the plot below - and it shows that we have good agreement, fun!
    \begin{center}
        \includegraphics[width=0.8\textwidth]{figures/best_fit.pdf}
    \end{center}
    So now I got intrigued and wondered what this meant for the mass of the planet. Since we know that $k_*$ is defined as
    \begin{equation}
        k_* = \qty(\frac{2 \pi G}{P (1 - e^2)})^{1/3} \frac{M_p \sin \iota}{(M_* + M_p)^{2/3}},
    \end{equation}
    where $\iota$ is the inclination and $M_*$ is the star mass. It should be fairly reasonable to rearrange this to find the planet mass and make the assumption that $M_* \gg M_p$ (this worries me slightly but I figure it'll be fine except in extreme cases). This gives that the planet mass is
    \begin{equation}
        M_p = \frac{k_* M_*^{2/3}}{\sin \iota} \qty(\frac{2 \pi G}{P (1 - e^2)})^{-1/3},
    \end{equation}
    So now we know everything in this equation from the best fit except $\iota$ and $M_*$. I made a grid of these values and plotted the planet mass as a contour plot below along with some lines indicating Earth and Neptune.

@@

    After some quick googling of the parameters that I found, it looks like our mystery planet is none other than HD 80606b! In this case, I can use the values of $M_* = 0.98 \pm 0.1 \unit{M_\odot}$ and $\iota = 89.6 \pm 0.4$ from \href{https://arxiv.org/abs/0902.4457}{Table 1 of this paper}. Plugging these into the equation above give that the mass of the planet should be
    \begin{equation}
        M_p \approx 5.59 \unit{M_{\rm Jup}}
    \end{equation}
    The real answer is $4.0 \pm 0.3 \unit{M_{\rm Jup}}$...but that's not far off given this is the most basic fit I could come up with in just a couple of hours, very fun! :D
}
